Road Segment Prioritization for Bicycle Infrastructure

### 2020-11-20 README.Rmd
updated draft paper

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. Therefore, the aim of this study is to develop a methodology that determines where cycling infrastructure should be added by accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand in a city will come from, how such demand should be routed on the street network, and how to  consider fair distribution of cycling infrastructure that is in line with egalitarian principles. The city of Manchester is used as a case study, but the methodology is generic enough to be applicable to any city. For routing, it is found that deviating from shortest paths to avoid high-stress road segments can result in acceptable travel time increases for cyclists. Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure. The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one. \\par\\textbf{Keywords:} cycling network, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

- Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand / (b) Routing / (c) Road Segment Prioritization
- Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them
- Should we use other cities for the sake of comparison (compare communities, road types etc)
- How to add tables properly (they are now added using latex code which only shows up when outputting to pdf)
- Footnotes are sometimes split between pages. How do I keep them on the same page?
- What is the best way to write pseudocode (Steps for Algorithm 1 and 2)

@@

All of these network-level methodologies are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors. This is important since different ethical principles constitute different problem formulations and targets. Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles. The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport]. @nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals. @lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area. @pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning. They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility. This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling. **Write some more here**

@@

Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network. To avoid this issue, @duthie2014optimization choose to ignore demand completely, and focus on creating a network that connects the entire study area. @olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.

@@

The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms)).

@@

Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity). When routing using the weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%. When routing on primary and trunk roads is banned (weighted_2 profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.

@@

One use-case of such granular data would be to identify roads that serve schools. The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years [@departmentgearchange2020], and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle. This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school [@departmentcycleinfradesign2020]. @goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%.

In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school. Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.

@@

After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments. This cumulative demand (referred to as *flow* is then used as a basis for determining where best to invest in segregated cycling infrastructure. In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. Therefore, the aim of this study is to develop a methodology that determines where cycling infrastructure should be added by accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand in a city will come from, how such demand should be routed on the street network, and how to  consider fair distribution of cycling infrastructure that is in line with egalitarian principles. The city of Manchester is used as a case study, but the methodology is generic enough to be applicable to any city. For routing, it is found that deviating from shortest paths to avoid high-stress road segments can result in acceptable travel time increases for cyclists. Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure. The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one, demostrating it's potential suitability in the planning of cycling networks. \\par\\textbf{Keywords:} cycling networks, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

- Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization
- Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them
- Should we use other cities for the sake of comparison (compare communities, road types etc)
- How to add tables properly (they are now added using latex code which only shows up when outputting to pdf)
- Footnotes are sometimes split between pages. How do I keep them on the same page?
- What is the best way to write pseudocode (Steps for Algorithm 1 and 2)

@@

All of these network-level methodologies are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors. This is important since different ethical principles constitute different problem formulations and targets. Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles. The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport]. @nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals. @lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area. @pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning. They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility. This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling. **Write some more here?**

@@

Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network. To avoid this issue, @duthie2014optimization choose to ignore existing demand completely, and focus on creating a network that connects the entire study area. @olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.

@@

The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms)).

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# this doesn't work because R project for code is in a different path to R project for paper (relative paths inside code won't work)
# source("R/_1.0_get_flow_data.R")
```

@@

Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity). When routing using the *weighted* weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%. When routing on primary and trunk roads is banned (*weighted_2* profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.

@@

One use-case of such granular data would be to identify roads that serve schools. The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years [@departmentgearchange2020], and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle. This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school [@departmentcycleinfradesign2020]. @goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%. In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school. Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.

@@

After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments. This cumulative demand (referred to as *flow*) is then used as a basis for determining where best to invest in segregated cycling infrastructure. In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.
